On the error performance of the $A_n$ lattices

### 2012-04-29 double.tex
minor edits?
Added to double

@@

We now state our main result.  The moment $\calM_n(m)$ of the lattice $A_n$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{H(n,m,k,a,b)},
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{(-1)^{k-a}2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1},
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$ (see Appendix~\ref{sec:solv-this-recurs}).  The first three such formula are:
\begin{align*}
\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of~}\vor(A_n),\\
\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the 2nd moment~ \cite[p. 462]{SPLAG}}, \\
\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}.
%\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
%\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.

@@

\begin{biography}{Robby G. McKilliam}
was born in Brisbane, Queensland, Australia in 1983. He received the B.Sc. degree in mathematics and the B.E. degree (Hons. I) in computer systems engineering from the University of Queensland, Brisbane, Australia in 2006.  In December 2010 he completed his PhD at the University of Queensland, Brisbane, Australia.  He is now at the Institute for Telecommunications Research at the University of South Australia, Adelaide.  His fields of interest are lattice theory, number theory, estimation theory and signal processing and communications.
\end{biography}

\begin{biography}{Ramanan Subramanian}
\end{biography}

\begin{biography}{Emanuele Viterbo}
received his degree (Laurea) in Electrical Engineering
in 1989 and his Ph.D. in 1995 in Electrical Engineering, both from the
Politecnico di Torino, Torino, Italy.  From 1990 to 1992 he was with the
European Patent Office, The Hague, The Netherlands, as a patent examiner
in the field of dynamic recording and error-control coding. Between 1995
and 1997 he held a post-doctoral position in the Dipartimento di
Elettronica of the Politecnico di Torino. In 1997-98 he was a
post-doctoral research fellow in the Information Sciences Research
Center of AT\&T Research, Florham Park, NJ, USA. He became first
Assistant Professor (1998) then Associate Professor (2005) in
Dipartimento di Elettronica at Politecnico di Torino. In 2006 he became
Full Professor in DEIS at University of Calabria, Italy. From 2010 he is
Full Professor in the ECSE Department and Associate Dean for Research
Training in the Faculty of Engineering at Monash University, Melbourne,
Australia.
Prof. Emanuele Viterbo is a 2011 Fellow of the IEEE,  a ISI Highly Cited
Researcher and Member of the Board of Governors of the IEEE Information
Theory Society (2011-2013). He is Associate Editor of IEEE Transactions
on Information Theory, European Transactions on Telecommunications and
Journal of Communications and Networks, and Guest Editor for IEEE
Journal of Selected Topics in Signal Processing: Special Issue Managing
Complexity in Multiuser MIMO Systems.
In 1993 he was visiting researcher in the Communications Department of
DLR, Oberpfaffenhofen, Germany. In 1994 and 1995 he was visiting the
cole Nationale Suprieure des Telcommunications (E.N.S.T.), Paris. In
2003 he was visiting researcher at the Maths Department of EPFL,
Lausanne, Switzerland. In 2004 he was visiting researcher at the
Telecommunications Department of UNICAMP, Campinas, Brazil. In 2005,
2006 and 2009 he was visiting researcher at the ITR of UniSA, Adelaide,
Australia. In 2007 he was visiting fellow at the Nokia Research Center,
Helsinki, Finland.
Dr. Emanuele Viterbo was awarded a NATO Advanced Fellowship in 1997 from
the Italian National Research Council. His main research interests are
in lattice codes for the Gaussian and fading channels, algebraic coding
theory, algebraic space-time coding, digital terrestrial television
broadcasting, and digital magnetic recording.
\end{biography}

\begin{biography}{I. Vaughan L. Clarkson}
was born in Brisbane,
Queensland, Australia, in 1968. He received the
B.Sc. degree in mathematics and the B.E. degree
(Hons. I) in computer systems engineering from
The University of Queensland, Brisbane, in 1989
and 1990, respectively, and the Ph.D. degree in
systems engineering from The Australian National
University, Canberra, in 1997.
In 1988, he was with the Defence Science and
Technology Organisation, Adelaide, Australia, ﬁrst
as a Cadet, later as a Professional Ofﬁcer, and ﬁnally
as a Research Scientist. From 1998 to 2000, he was a Lecturer at The University
of Melbourne, Melbourne, Australia. From 2000 to 2008, he was a Senior
Lecturer in the School of Information Technology and Electrical Engineering
at The University of Queensland. In 2008, he was promoted to Reader. In 2005,
he was a Visiting Professor in the Department of Electrical and Computer
Engineering, The University of British Columbia, Vancouver, Canada. His
research interests include statistical signal processing for communications and
defense, image processing, information theory, and lattice theory.
\end{biography}

>>>>

\begin{figure*}[bt]
% ensure that we have normalsize text
%\normalsize
\small
% Store the current equation number.
%\setcounter{MYtempeqncnt}{\value{equation}}
% Set the equation number to one less than the one
% desired for the first equation here.
% The value here will have to changed if equations
% are added or removed prior to the place these
% equations are referenced in the main text.
%\setcounter{equation}{5}
\begin{align*}
\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}} \\
\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}} \\
\calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6+341550 n^7+75526 n^8+8855 n^9+385 n^{10}}{95800320 (1 + n)^{9/2}}
\end{align*}
% Restore the current equation number.
%\setcounter{equation}{\value{MYtempeqncnt}}
% IEEE uses as a separator
\hrulefill
% The spacer can be tweaked to stop underfull vboxes.
\vspace*{4pt}
\end{figure*}

@@

We now state our main result.  The moment $\calM_n(m)$ of the lattice $A_n$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{H(n,m,k,a,b)},
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{(-1)^{k-a}2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1},
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$ (see Appendix~\ref{sec:solv-this-recurs}).  The first three such formula are:
\begin{align*}
\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of~}\vor(A_n),\\
\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the 2nd moment~ \cite[p. 462]{SPLAG}}, \\
\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}
%\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
%\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
and the next three formula are displayed at the top of this page.  We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.

@@

\begin{IEEEbiographynophoto}{Robby G. McKilliam}
was born in Brisbane, Queensland, Australia in 1983. He received the B.Sc. degree in mathematics and the B.E. degree (Hons. I) in computer systems engineering from the University of Queensland, Brisbane, Australia in 2006.  In December 2010 he completed his PhD at the University of Queensland, Brisbane, Australia.  He is now at the Institute for Telecommunications Research at the University of South Australia, Adelaide.  His fields of interest are lattice theory, number theory, estimation theory and signal processing and communications.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Ramanan Subramanian}
received the Bachelor of Technology degree in Electrical Engineering from the Indian Institute of Technology, Madras in 2003, the Master of Science and PhD degrees in Electrical and Computer Engineering from the Georgia Institute of Technology, Atlanta, USA in 2006 and 2009 respectively. Since late 2009, he has been working as a research fellow at the Institute for Telecommunications Research (ITR) at the University of South Australia in Adelaide. His research interests include wireless networking, information theory, queuing theory, and statistics.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Emanuele Viterbo}
received his degree (Laurea) in Electrical Engineering
in 1989 and his Ph.D. in 1995 in Electrical Engineering, both from the
Politecnico di Torino, Torino, Italy.  From 1990 to 1992 he was with the
European Patent Office, The Hague, The Netherlands, as a patent examiner
in the field of dynamic recording and error-control coding. Between 1995
and 1997 he held a post-doctoral position in the Dipartimento di
Elettronica of the Politecnico di Torino. In 1997-98 he was a
post-doctoral research fellow in the Information Sciences Research
Center of AT\&T Research, Florham Park, NJ, USA. He became first
Assistant Professor (1998) then Associate Professor (2005) in
Dipartimento di Elettronica at Politecnico di Torino. In 2006 he became
Full Professor in DEIS at University of Calabria, Italy. From 2010 he is
Full Professor in the ECSE Department and Associate Dean for Research
Training in the Faculty of Engineering at Monash University, Melbourne,
Australia.
Prof. Emanuele Viterbo is a 2011 Fellow of the IEEE,  a ISI Highly Cited
Researcher and Member of the Board of Governors of the IEEE Information
Theory Society (2011-2013). He is Associate Editor of IEEE Transactions
on Information Theory, European Transactions on Telecommunications and
Journal of Communications and Networks, and Guest Editor for IEEE
Journal of Selected Topics in Signal Processing: Special Issue Managing
Complexity in Multiuser MIMO Systems.
In 1993 he was visiting researcher in the Communications Department of
DLR, Oberpfaffenhofen, Germany. In 1994 and 1995 he was visiting the
cole Nationale Suprieure des Telcommunications (E.N.S.T.), Paris. In
2003 he was visiting researcher at the Maths Department of EPFL,
Lausanne, Switzerland. In 2004 he was visiting researcher at the
Telecommunications Department of UNICAMP, Campinas, Brazil. In 2005,
2006 and 2009 he was visiting researcher at the ITR of UniSA, Adelaide,
Australia. In 2007 he was visiting fellow at the Nokia Research Center,
Helsinki, Finland.
Dr. Emanuele Viterbo was awarded a NATO Advanced Fellowship in 1997 from
the Italian National Research Council. His main research interests are
in lattice codes for the Gaussian and fading channels, algebraic coding
theory, algebraic space-time coding, digital terrestrial television
broadcasting, and digital magnetic recording.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{I. Vaughan L. Clarkson}
was born in Brisbane,
Queensland, Australia, in 1968. He received the
B.Sc. degree in mathematics and the B.E. degree
(Hons. I) in computer systems engineering from
The University of Queensland, Brisbane, in 1989
and 1990, respectively, and the Ph.D. degree in
systems engineering from The Australian National
University, Canberra, in 1997.
In 1988, he was with the Defence Science and
Technology Organisation, Adelaide, Australia, ﬁrst
as a Cadet, later as a Professional Ofﬁcer, and ﬁnally
as a Research Scientist. From 1998 to 2000, he was a Lecturer at The University
of Melbourne, Melbourne, Australia. From 2000 to 2008, he was a Senior
Lecturer in the School of Information Technology and Electrical Engineering
at The University of Queensland. In 2008, he was promoted to Reader. In 2005,
he was a Visiting Professor in the Department of Electrical and Computer
Engineering, The University of British Columbia, Vancouver, Canada. His
research interests include statistical signal processing for communications and
defense, image processing, information theory, and lattice theory.
\end{IEEEbiographynophoto}
